Lesson 4
========================================================

***

### Scatterplots and Perceived Audience Size

### Scatterplots

```{r Scatterplots}
library('ggplot2')
pf <- read.csv('./pseudo_facebook.tsv', sep = '\t')

# Scatterplot
qplot(data = pf, x = age, y = friend_count)
```

***

#### What are some things that you notice right away?
Response: Younger people have more friends than the others

***

### ggplot Syntax

```{r ggplot Syntax}
ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(na.rm = TRUE) + 
  scale_x_continuous(breaks = seq(13,90,5), limits = c(13,90))
```

***

### Overplotting
Notes: Bottom part of the plot has too many points which makes it harder to understand

```{r Overplotting}
# Adding Transparency
# 20 points makes one black point
ggplot(data = pf, aes(x = age, y = friend_count)) +
  #geom_point(alpha = 1/20, na.rm = TRUE) +
  geom_jitter(alpha = 1/20, na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,5), limits = c(13,90))
```

#### What do you notice in the plot?
Response: Most people have around 1000-2000 friends
***

### Coord_trans()
Notes:

```{r Coord_trans()}
ggplot(data = pf, aes(x = age, y = friend_count + 1)) +
  geom_point(alpha = 1/20, na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,5), limits = c(13,90)) +
  coord_trans(y = 'sqrt') + 
  ylab('Friend Count')
```

***

### Alpha and Jitter
Notes:


```{r}
ggplot(data = pf, aes(x = age, y = pf$friendships_initiated)) +
  geom_jitter(alpha = 1/10, na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,10), limits = c(13,90)) +
  scale_y_continuous(limits = c(0,2000), breaks = seq(0,2000,500)) +
  ylab('Friendships Initiated')

# Alternate Syntax
ggplot(data = pf, aes(x = age, y = pf$friendships_initiated)) +
  geom_point(alpha = 1/10, position = 'jitter', na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,10), limits = c(13,90)) +
  scale_y_continuous(limits = c(0,2000), breaks = seq(0,2000,500)) +
  ylab('Friendships Initiated')
```

***

### Overplotting and Domain Knowledge
Notes: Used Percentage for Axes to study the data
***

### Conditional Means

```{r Conditional Means}
library(dplyr)
# age_by_group <- group_by(pf, age)
# pf.fc_by_age <- summarise(
#   age_by_group,
#   friend_count_mean = mean(friend_count),
#   friend_count_median = median(friend_count),
#   n = n()
# )
# 
# pf.fc_by_age <-  arrange(pf.fc_by_age)

# Using pipes in R
pf.fc_by_age <- pf %>% 
                group_by(age) %>% 
                summarise(
                friend_count_mean = mean(friend_count),
                friend_count_median = median(friend_count),
                n = n()) %>%  
                arrange()

head(pf.fc_by_age)
```


```{r Conditional Means Plot}
ggplot(data = pf.fc_by_age, 
       mapping = aes(x = age, friend_count_mean)) +
  geom_point(na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,10), limits = c(13,90)) +
  scale_y_continuous(breaks = seq(0,450,50), limits = c(0,450))

# Using geom_line() to see a trend

ggplot(data = pf.fc_by_age, 
       mapping = aes(x = age, friend_count_mean)) +
  geom_line(na.rm = TRUE) +
  scale_x_continuous(breaks = seq(13,90,10), limits = c(13,90)) +
  scale_y_continuous(breaks = seq(0,450,50), limits = c(0,450))
```

***

### Overlaying Summaries with Raw Data

```{r Overlaying Summaries with Raw Data, warning=FALSE}
ggplot(data = pf, aes(x = age, y = friend_count + 1)) +
  geom_point(alpha = 1/20, color = 'orange') +
  scale_x_continuous(breaks = seq(13,90,5), limits = c(13,90)) +
  coord_trans(y = 'sqrt') + 
  geom_line(stat = 'summary', fun.y = mean) +
  geom_line(stat = 'summary', fun.y = median) +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .9), color = 'red', linetype = 2) +
  geom_line(stat = 'summary', fun.y = quantile,
            fun.args = list(probs = .1), color = 'red', linetype = 2) +
  ylab('Friend Count')
```

### Correlation
Notes:

```{r Correlation}

```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
# Checking correlation
#cor.test(pf$age, pf$friend_count)
with(pf, cor.test(age, friend_count))
```

***

### Correlation Methods
Notes: Use Pearson by default

***

## Create Scatterplots

```{r}
ggplot(data = pf, mapping = 
         aes(x = likes_received, y = www_likes_received)) +
  geom_point()
```

***

### Strong Correlations
Notes:

```{r Strong Correlations}
ggplot(data = pf, mapping = 
         aes(x = likes_received, y = www_likes_received)) +
  geom_point(alpha = 1/20, na.rm = TRUE) +
  xlim(0, quantile(pf$likes_received, 0.95)) +
  ylim(0, quantile(pf$www_likes_received, 0.95)) +
  geom_smooth(method = 'lm', color = 'red')
```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}
cor.test(pf$likes_received, pf$www_likes_received)
```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
#install.packages('alr3')
library(alr3)
data("Mitchell")
summary(Mitchell)
```

Create your plot!

```{r Temp vs Month}
ggplot(data = Mitchell, mapping = aes(x = Month, y = Temp)) +
  geom_point()
```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot. 0.01

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}
cor.test(Mitchell$Month, Mitchell$Temp)
```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}
ggplot(data = Mitchell, mapping = aes(x = Month, y = Temp)) +
  geom_point(color = 'red') + 
  scale_x_continuous(breaks = seq(0,203,12), 
                     limits = c(0,205))
```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

```

***

### Age with Months Means

```{r Age with Months Means}

```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

