---
title: "Transforming Data in R"
author: "Pramod Duvvuri"
date: "3/29/2019"
output:
  pdf_document: 
    fig_caption: yes
  word_document: default
---

## Data Manipulation in R

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=5)
```

```{r}
library(nycflights13) # Data to be used
library(tidyverse)
```

```{r}
nycflights13::flights
attach(flights)
?flights
```

### Basics of dplyr

The below functions will be covered in this file.
```{r, eval=FALSE}
1. filter()
2. arrange()
3. select()
4. mutate()
5. summarise()
6. group_by()
```

### filter()

```{r}
# Select all flights on January 1st
filter(flights, month == 1, day == 1)
```

```{r}
jan_1_data <- filter(flights, month == 1, day == 1) # Can use view(jan_1_data) to see the data

# To assing the data to another variable and print to console
(dec25_data <- filter(flights, month == 12, day == 25))
```

```{r, eval=FALSE}
filter(flights, month = 4) # throws an error (must use ==)
```

```{r}
sqrt(2) ^ 2 == 2
1 / 49 * 49 == 1
```

```{r}
# Use near() for approximation
near(sqrt(2) ^ 2, 2)
near(1 / 49 * 49, 1)
```

```{r}
# Using filter() with logical operators
filter(flights, month == 4 | month == 6)
# Alternate way to write the above line of code
apr_jun_data <- filter(flights, month %in% c(4,7))
nrow(apr_jun_data)
```

```{r}
# Missing Values
NA == NA
# Adding some context
x <- NA # Age of Person 1
y <- NA # Age of Person 2
x == y # Compare
is.na(x)
```

```{r}
# Handling Missing Values using filter()
df <- tibble(x = c(19,27,32,NA))
filter(df, x > 1) # NA excluded
filter(df, is.na(x) | x > 1)
```

### Exercises

Q1. Find all flights that

Q1.1 Had an arrival delay of two or more hours
```{r}
filter(flights, arr_delay >= 120)
```


Q1.2 Flew to Houston (IAH or HOU)
```{r}
filter(flights, dest %in% c('IAH','HOU'))
```


Q1.3 Were operated by United, American, or Delta
```{r}
filter(flights, carrier %in% c('AA','DL','UA'))
```


Q1.4 Departed in summer (July, August, and September)
```{r}
filter(flights, month %in% c(7,8,9))
```


Q1.5 Arrived more than two hours late, but didn’t leave late
```{r}
filter(flights, arr_delay > 120, dep_delay <= 0)
```


Q1.6 Were delayed by at least an hour, but made up over 30 minutes in flight
```{r}
filter(flights, dep_delay > 60, arr_delay < 30)
```


Q1.7 Departed between midnight and 6am (inclusive)
```{r}
filter(flights,dep_time >= 0000 & dep_time <= 0600)
```


```{r}
# Using between() for filtering
sum(between(flights$dep_time, 0000, 0600), na.rm = TRUE)
```

### arrange()

```{r}
arrange(flights, year, month, day)
```

```{r}
# Descending Order
arrange(flights, desc(dep_delay))
```

```{r}
# Missing Values (stored in the end)
df <- tibble(x = c(5, 2, NA))
arrange(df, x)
arrange(df, desc(x))
```

### Exercises


Q1. How could you use arrange() to sort all missing values to the start? (Hint: use is.na()).
```{r}
arrange(df, desc(is.na(x)))
```


Q2. Sort flights to find the most delayed flights. Find the flights that left earliest.
```{r}
arrange(flights, desc(arr_delay), dep_delay)
```


Q3. Sort flights to find the fastest flights.
```{r}
arrange(flights, arr_delay)
```

Q4. Which flights travelled the longest? Which travelled the shortest?
```{r}
arrange(flights, air_time) # shortest flights
arrange(flights, desc(air_time)) # longest flights
```


### select()

```{r}
# Used to select subset of columns/features
select(flights,year,month,day)
# Alteranate way to get same subset of data
select(flights, year:day)
# All columns but a few (omit a few columns)
select(flights, -(year:day))
select(flights, -c(year,month,day))
select(flights,-year,-month,-day)
```

### rename()

```{r}
# Renaming variables using rename()
rename(flights, tail_number = tailnum)
```

### everything()

```{r}
# everything()
select(flights, air_time, everything()) # moves the air_time column to the beginning
```

### Exercises

Q1. Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights
```{r}
select(flights, c(dep_time, dep_delay, arr_time, arr_delay))
```


Q2. What happens if you include the name of a variable multiple times in a select() call?
```{r}
select(flights, c("arr_time","arr_time")) # no error
```


Q3. What does the one_of() function do? Why might it be helpful in conjunction with this vector?
```{r}
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
select(flights, one_of(vars))
```


Q4. Does the result of running the following code surprise you? How do the select helpers deal with case by default? How can you change that default?
```{r}
select(flights, contains("TIME")) # ignore.case can be set to FALSE (default TRUE)
```

### mutate()

```{r}
# Using subset of flight data
flights_sml <- select(flights, 
  year:day, 
  ends_with("delay"), 
  distance, 
  air_time
)

# Adding new columns 

mutate(flights_sml,
  gain = dep_delay - arr_delay,
  speed = round(distance / air_time * 60,3)
)
```

```{r}
# Use columns just created
mutate(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours # gain created above
)
```

### transmute()

```{r}
# Keep only newly created columns
transmute(flights_sml,
  gain = dep_delay - arr_delay,
  hours = air_time / 60,
  gain_per_hour = gain / hours
)
```

### Aggregate Functions 

```{r}
y <- c(1, 2, 2, NA, 3, 4)
min_rank(y)
row_number(y)
dense_rank(y)
```

### Exercises

```{r}
transmute(flights, air_time, arr_time - dep_time)
#transmute(flights, arr_time, sched_arr_time, arr_delay, dep_time, sched_dep_time, dep_delay)
```

```{r}
transmute(flights, dep_time, sched_dep_time, dep_delay)
```

```{r}
fl_df <- mutate(flights, total_delay = arr_delay + dep_delay)
transmute(arrange(fl_df, desc(total_delay)), total_delay)
```

```{r, eval=FALSE}
1:3 + 1:10 # error
```

### summarise()

```{r}
# One row summary
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

### group_by()

```{r}
# Grouping and applying summarise()
by_day <- group_by(flights, year, month, day)
summarise(by_day, delay = round(mean(dep_delay, na.rm = TRUE),2))
```


```{r}
by_dest <- group_by(flights, dest)

delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delay <- filter(delay, count > 20, dest != "HNL")
```


```{r}
ggplot(data = delay, mapping = aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```

### Multiple Operations using pipes

```{r}
delays <- flights %>% 
  group_by(dest) %>% 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) %>% 
  filter(count > 20, dest != "HNL")
```

```{r}
# na.rm = TRUE removes missing values
flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay))

flights %>% 
  group_by(year, month, day) %>% 
  summarise(mean = mean(dep_delay, na.rm = TRUE))
```

```{r}
# remove flights with NA values before exploring dataset

not_cancelled <- flights %>% 
  filter(!is.na(dep_delay), !is.na(arr_delay))

# Cancelled Flights
print(nrow(flights) - nrow(not_cancelled))
```

```{r}
not_cancelled %>% 
  group_by(year,month) %>% 
  summarise(mean_delay = round(mean(dep_delay),2)) %>% arrange(mean_delay)
```

```{r}
delays <- not_cancelled %>% 
  group_by(tailnum) %>% 
  summarise(
    delay = mean(arr_delay)
  )

ggplot(data = delays, mapping = aes(x = delay)) + 
  geom_freqpoly(binwidth = 10)
```


