---
title: "Exploratory Data Analysis in R"
author: "Pramod Duvvuri"
date: "3/30/2019"
output: pdf_document
---

## Introduction to EDA

The common steps in Exploratory Data Analysis (EDA) are:

1. Generate questions about your data.

2. Search for answers by visualising, transforming, and modelling your data.

3. Use what you learn to refine your questions and/or generate new questions.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=5)
```

```{r echo=TRUE}
library(tidyverse)
library(nycflights13)
```

## Sample Visualizations

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

```{r}
# Heights of bars using dplyr::count()
diamonds %>% count(cut)
```

```{r}
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = carat), binwidth = 0.5)
```

```{r}
# Heights using dplyr and ggplot2
diamonds %>% count(cut_width(carat,0.5))
```

```{r}
# Multiple histograms ovrlapping in a single plot
ggplot(data = filter(diamonds, carat < 3), mapping = aes(x = carat, colour = cut)) +
  geom_freqpoly(binwidth = 0.1)
```

## Sample Questions

1. Which values are the most common? Why?

2. Which values are rare? Why? Does that match your expectations?

3. Can you see any unusual patterns? What might explain them?


```{r}
ggplot(data = faithful, mapping = aes(x = eruptions)) + 
  geom_histogram(binwidth = 0.25) + ggtitle('Frequency of Length of Eruptions')
```


## Handling Data

1. Typical Values
2. Unusual Values
3. Missing Values

### Unusual Values

```{r}
summary(diamonds)
```

```{r}
ggplot(data = diamonds) + geom_histogram(mapping = aes(x = y), binwidth = 0.5)
```

```{r}
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = y), binwidth = 0.5) +
  coord_cartesian(ylim = c(0,100))
```

```{r}
# Digging unusual values using dplyr
unusual <- diamonds %>% 
  filter(y < 3 | y > 20) %>%
  select(price, x , y , z) %>%
  arrange(y)

unusual
```

### Missing Values

We shall replace unusual values in the data as missing values (NA) instead of dropping them

#### ifelse()

```{r}
# case_when() can also be used to re-write the below line of code
diamonds2 <- diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y))
```

```{r}
ggplot(data = diamonds2, mapping = aes(x = x, y = y)) +
  geom_point(na.rm = TRUE)
```

```{r}
# Compare cancelled and non-cancelled flights
nycflights13::flights %>% 
  mutate(
    cancelled = is.na(dep_time),
    sched_hour = sched_dep_time %/% 100,
    sched_min = sched_dep_time %% 100,
    sched_dep_time = sched_hour + sched_min / 60
  ) %>% 
  ggplot(mapping = aes(sched_dep_time)) + 
    geom_freqpoly(mapping = aes(colour = cancelled), binwidth = 1/4)
```

```{r}
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x = cut ,y = price))
```

### reorder()

```{r}
# Reordering basing on average price from lower to higher
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x = reorder(cut, price, FUN = mean), y = price))
```

```{r}
# Flipping can be done if variable names are long
ggplot(data = diamonds) + 
  geom_boxplot(mapping = aes(x = reorder(cut, price, FUN = mean), y = price)) +
  coord_flip()
```