---
title: "Introduction to Visualization in R"
author: "Pramod Duvvuri"
date: "3/22/2019"
output:
  pdf_document: 
    fig_caption: yes
  word_document: default
---

## Introduction to ggplot2

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width=9, fig.height=5)
```

```{r render=FALSE}
library(tidyverse)
```

```{r}
mpg
attach(mpg)
```

```{r eval=FALSE}
# ggplot2 TEMPLATE
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```

```{r}
# Add data points
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

### Exercises-1

```{r}
# Question-1
ggplot(data = mpg)
```

```{r}
# Question-2
nrow(mpg) # Rows
ncol(mpg) # Columns
```

```{r eval=FALSE}
# Question-3
?mpg
# Answer: f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```

```{r}
# Question-4
ggplot(data = mpg) + geom_point(mapping = aes(x = cyl, y = hwy))
```

```{r}
# Question-5
ggplot(data = mpg) + geom_point(mapping = aes(x = class, y = drv))
```

## Aesthetic Mappings

```{r}
# Adding aesthetic to the plot (color/size/shape could be used here)
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy,
                                              color = class))
```

```{r}
# Using size as aesthetic instead of color
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy,
                                              size = class))
```

```{r}
# Alpha and Shape Aesthetics
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy,
                                              alpha = class))

ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy,
                                              shape = class))
```


```{r}
# Using a different color for the data points
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy), 
                                              color = "blue") 
```

### Exercises-2

Q1. What’s gone wrong with this code? Why are the points not blue? 
```{r}
# color should be outside aes()
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
```

Q2. Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
```{r}
summary(mpg)
```

Q3. Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
```{r}
# Mapping a continuos variable to color
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
                                             color = cyl))
```

```{r eval=FALSE}
# Shape cannot be mapped, the line below throws an error
#ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
#                                             shape = cty))
```

```{r}
# Mapping a continuos variable to size
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
                                             size = cyl))
```

```{r}
# Mapping same variable to multiple aesthetics (Discrete Variable)
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
                                             color = trans, size = trans))
```

```{r}
# Mapping same variable to multiple aesthetics (Continuos Variable)
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
                                             color = cty, size = cty))
```

Q. What does the stroke aesthetic do? What shapes does it work with? (Hint: use ?geom_point)
```{r}
?geom_point # stroke aesthetic
```

Q. What happens if you map an aesthetic to something other than a variable name, like aes(colour = displ < 5)? Note, you’ll also need to specify x and y. 
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy, 
                                             color = displ < 5))
```

## Facets

```{r}
# Original Plot
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
# Using facet_wrap()
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
# Using facet_grid()
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(drv ~ cyl)
```

```{r}
# Using facet_grid()
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_grid(. ~ cyl)
```

### Exercises-3

Q1. What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ cty, nrow = 2)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = drv, y = cyl))
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(drv ~ .)

ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  facet_grid(. ~ cyl)
```

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```

```{r}
?facet_wrap
```

## Geometric Objects

```{r}
# Using geom_smooth()
ggplot(data = mpg) + geom_smooth(mapping = 
                                   aes(x = displ, y = hwy))
```

```{r}
# Different linetype aesthetic
ggplot(data = mpg) + geom_smooth(mapping = 
                                   aes(x = displ, y = hwy, linetype = drv))
```

```{r}
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ, y = hwy, group = drv))
```

```{r}
# Disabling Legend
ggplot(data = mpg) + geom_smooth(mapping = aes(x = displ,
                                               y = hwy, color = drv),
                                               show.legend = FALSE)
```

```{r}
# Multiple geoms in a Plot
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) +
  geom_smooth(mapping = aes(x = displ, y = hwy))
```

```{r}
# Global Mappings
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point() + geom_smooth()
```

```{r}
# Local Mappings
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth()
```

### Exercises-4

Q1. What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?
Ans. geom_boxplot(), geom_histogram()
```{r}
ggplot(data = mpg, mapping = aes(x = drv, y = hwy)) + geom_boxplot()
```

Q2. Recreate the plots
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)
```

Q3. Recreate the plots
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
         geom_point() +
         geom_smooth(mapping = aes(group = drv), se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
         geom_point() +
         geom_smooth(mapping = aes(group = drv), se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) +
         geom_point() +
         geom_smooth(mapping = aes(group = drv, linetype = drv), se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
         geom_point(mapping = aes(color = drv)) +
         geom_smooth(se = FALSE)
```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) +
         geom_point(mapping = aes(color = drv))
```

## Statistical transformations

```{r}
# Bar Charts
# Using the diamonds dataset
attach(diamonds)
nrow(diamonds)
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut))
```

```{r}
# Using stat_count() instead of geom_bar()
ggplot(data = diamonds) + 
  stat_count(mapping = aes(x = cut))
```

Every geom function has a stat equivalent and vice-versa that could be used to produce the same plot in two different ways.

```{r}
# Displaying Proportion instead of frequency
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
```

```{r}
ggplot(data = diamonds) + 
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```

### Exercises-5

Q1. What is the default geom associated with stat_summary()? How could you rewrite the previous plot to use that geom function instead of the stat function?
```{r}
ggplot(data = diamonds) + geom_boxplot(mapping = aes(x = cut, y = depth))
```

## Position Adjustments

```{r}
# Using Colour aesthetic
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, colour = cut))
```

```{r}
# Using fill aesthetic
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = cut))
```

```{r}
# Stacked Bar Chart
ggplot(data = diamonds) + 
  geom_bar(mapping = aes(x = cut, fill = clarity))
```

### Exercises-6

Q.What is the problem with this plot? How could you improve it?
```{r}
# Original Plot
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

```{r}
# Adding jitter
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point(position = "jitter")
```

The above plot can also be plotted using the geom_jitter function. This jitter avoids overplotting and adds a small amount on random noise to each point.

Q. What’s the default position adjustment for geom_boxplot()? Create a visualisation of the mpg dataset that demonstrates it.
```{r}
ggplot(data = mpg, mapping = aes(x = drv, y = hwy)) + geom_boxplot()
```

```{r, eval=FALSE}
# Grammar of Graphics Template
ggplot(data = <DATA>) + 
  <GEOM_FUNCTION>(
     mapping = aes(<MAPPINGS>),
     stat = <STAT>, 
     position = <POSITION>
  ) +
  <COORDINATE_FUNCTION> +
  <FACET_FUNCTION>
```

