Lesson 3
===============================

***

### What to Do First?
Notes: What is the goal ? What questions are you trying to answer ?

***

### Pseudo-Facebook User Data
Notes: Loaded Data

```{r Pseudo-Facebook User Data}
pf <- read.csv('./pseudo_facebook.tsv', sep = '\t')
names(pf)
```

***

### Histogram of Users' Birthdays

```{r Histogram of Users\' Birthdays}
# install.packages('ggplot2')
library(ggplot2)
# install.packages('ggthemes', dependencies = TRUE)
library(ggthemes)
qplot(data = pf, x = dob_day)
qplot(data = pf, x = dob_day) +
  scale_x_continuous(breaks = 1:31)

```

***

#### What are some things that you notice about this histogram?
Response: Most people are born on the first day of a month. Less number of people are born on the 31st because not all months have 31 days.

***

### Moira's Investigation
Notes: Perception of audience on Facebook. How many people view your posts on Facebook ? A project was tackling this problem.

***

### Faceting
Notes:

```{r Faceting}
ggplot(data = pf, aes(x = dob_day)) +
  geom_histogram(binwidth = 1) +
  scale_x_continuous(breaks = 1:31) +
  facet_wrap(~dob_month)
```

#### Let’s take another look at our plot. What stands out to you here?
Response: More people have birthdays on the first of January

***

### Be Skeptical - Outliers and Anomalies
Notes: Plot and observe these anomalies and outliers

***

### Moira's Outlier
Notes:
#### Which case do you think applies to Moira’s outlier?
Response: Bad data
***

### Friend Count
Notes:

#### What code would you enter to create a histogram of friend counts?

```{r Friend Count}
ggplot(data = pf) + 
  geom_histogram(mapping = aes(x = friend_count))
```

#### How is this plot similar to Moira's first plot?
Response:

***

### Limiting the Axes
Notes:

```{r Limiting the Axes}
ggplot(data = pf, mapping = aes(x = friend_count)) + 
  geom_histogram() +
  scale_x_continuous(limits = c(0,1000), 
                     breaks = seq(0,1000,50)) + 
  stat_bin(binwidth = 25)
```

### Exploring with Bin Width

### Adjusting the Bin Width

### Faceting Friend Count
```{r Faceting Friend Count}
# What code would you add to create a facet the histogram by gender?
# Add it to the code below.
ggplot(data = pf, mapping = aes(x = friend_count)) + 
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender) + 
  coord_flip()
```

***

### Omitting NA Values
Notes:

```{r Omitting NA Values}
ggplot(data = subset(pf, !is.na(pf$gender)), 
       mapping = aes(x = friend_count)) + 
  geom_histogram() +
  scale_x_continuous(limits = c(0, 1000),
                     breaks = seq(0, 1000, 50)) + 
  facet_wrap(~gender) + 
  coord_flip()
```

***

### Statistics 'by' Gender
Notes:

```{r Statistics \'by\' Gender}
table(pf$gender)
by(pf$friend_count, pf$gender, summary)
```

#### Who on average has more friends: men or women?
Response: Women

#### What's the difference between the median friend count for women and men?
Response: 22

#### Why would the median be a better measure than the mean?
Response: Number of samples does not affect it

***

### Tenure
Notes:

```{r Tenure}
ggplot(aes(x = tenure), data = pf) +
  geom_histogram(binwidth = 30, 
                 color = 'black', fill = '#099DD9')
```

***

#### How would you create a histogram of tenure by year?

```{r Tenure Histogram by Year}
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(color = 'black', 
                 fill = '#099DD9', 
                 binwidth = 0.25)
```

***

### Labeling Plots
Notes:

```{r Labeling Plots}
# Histogram with more details and labels
ggplot(aes(x = tenure/365), data = pf) +
  geom_histogram(color = 'black', 
                 fill = '#099DD9', 
                 binwidth = 0.25, na.rm = TRUE) +
  scale_x_continuous(limits = c(0,7), 
                     breaks = seq(0,7,1)) +
  scale_y_continuous(limits = c(0,15000)) +
  xlab('Number of years using Facebook') +
  ylab('Number of users in sample')
```

***

### User Ages
Notes:

```{r User Ages}
ggplot(data = pf, mapping = aes(x = pf$age)) + 
  geom_histogram(binwidth = 5, na.rm = TRUE) + 
  scale_x_continuous(limits = c(13,120), breaks = seq(17,120,5)) +
  xlab('Age')
```
***

### The Spread of Memes
Notes: Log Scale was used since the count was in the high 1000's to identify patterns which were not visible on a normal scale

***

### Add a Scaling Layer
Notes:

```{r Add a Scaling Layer}
#install.packages('gridExtra')
library(gridExtra)
# Normal Histogram

plot_1 <- ggplot(data = pf, mapping = aes(x = friend_count + 1)) + 
  geom_histogram(na.rm = TRUE) + xlab('Number of Friends on Facebook')

# Histogram using log10 transformation
plot_2 <- plot_1 + 
  scale_x_log10()

# Histogram using sqrt transformation
plot_3 <- plot_1 + 
  scale_x_sqrt(limits = c(1,1000))

grid.arrange(plot_1, plot_2, plot_3)
```

***


### Frequency Polygons

```{r Frequency Polygons}
ggplot(aes(x = friend_count + 1, y = ..count../sum(..count..)),
       data = subset(pf, !is.na(gender))) +
  geom_freqpoly(na.rm = TRUE ,aes(color = gender), binwidth=10) +
  scale_x_continuous(limits = c(0, 1000), breaks = seq(0, 1000, 50)) +
  xlab('Friend Count') +
  ylab('Proportion of users with that friend count')
```

***

### Likes on the Web
Notes:

```{r Likes on the Web}
ggplot(aes(x = www_likes), data = subset(pf, !is.na(gender))) +
  geom_freqpoly(aes(color = gender)) + 
  scale_x_continuous() +
  scale_x_log10() +
  xlab('Likes')
```


***

### Box Plots
Notes:

```{r Box Plots}
ggplot(data = subset(pf, !is.na(pf$gender)), mapping = aes(x = gender,y = friend_count)) + 
  geom_boxplot()
```

#### Adjust the code to focus on users who have friend counts between 0 and 1000.

```{r}
ggplot(data = subset(pf, !is.na(pf$gender)), 
       mapping = aes(x = gender,y = friend_count)) + 
  geom_boxplot(na.rm = TRUE) +
  scale_y_continuous(limits = c(0,1000))

# Using coord_cartesian()

ggplot(data = subset(pf, !is.na(pf$gender)), 
       mapping = aes(x = gender,y = friend_count)) + 
  geom_boxplot(na.rm = TRUE) +
  coord_cartesian(ylim = c(0,1000))
```

***

### Box Plots, Quartiles, and Friendships
Notes:

```{r Box Plots, Quartiles, and Friendships}
ggplot(data = subset(pf, !is.na(pf$gender)), 
       mapping = aes(x = gender,y = friend_count)) + 
  geom_boxplot(na.rm = TRUE) +
  coord_cartesian(ylim = c(0,250))

by(pf$friend_count, pf$gender, summary)

ggplot(data = subset(pf, !is.na(pf$gender)), 
       mapping = aes(x = gender,y = friendships_initiated)) + 
  geom_boxplot(na.rm = TRUE) +
  coord_cartesian(ylim = c(0,150))
```

#### On average, who initiated more friendships in our sample: men or women?
Response: From plot 3, women have initiated more friendships by just a small margin
#### Write about some ways that you can verify your answer.
Response: The boxplot can be used to verify the same

```{r Friend Requests by Gender}
by(pf$friendships_initiated, pf$gender, summary)
```

Response: Female have initiated more friendships

***

### Getting Logical
Notes:

```{r Getting Logical}
summary(pf$mobile_likes)
summary(pf$mobile_likes > 0)
mobile_check_in <- NA
pf$mobile_check_in <- ifelse(pf$mobile_likes > 0, 1, 0)
pf$mobile_check_in <- factor(pf$mobile_check_in)
# Calculating Percentage
round(summary(pf$mobile_check_in)/ length(pf$mobile_check_in) * 100,2)
```

Response: More people use mobile check in according to this sample we have at hand

***