---
output:
  html_document: default
---
Lesson 5
========================================================

### Multivariate Data
***

***

### Third Qualitative Variable

```{r Third Qualitative Variable}
library(tidyverse)
pf <- read.delim('pseudo_facebook.tsv', sep = '\t')

ggplot(aes(x = age),
       data = subset(pf, !is.na(gender))) + 
  geom_histogram(na.rm = TRUE, bins = 50) + 
  facet_wrap(~gender)
```

***

### Plotting Conditional Summaries

```{r Manipulating Data}
pf.fc_by_age_gender <- pf %>% 
                       filter(!is.na(gender)) %>% 
                       group_by(age, gender) %>% 
                       summarise(mean_friend_count = mean(friend_count),
                                 median_friend_count = median(friend_count),
                                 n = n()) %>% 
                       ungroup() %>% 
                       arrange(age)

head(pf.fc_by_age_gender)
```

```{r Plotting Conditional Summaries}
ggplot(aes(x = age, y = median_friend_count),
       data = pf.fc_by_age_gender) +
  geom_line(aes(color = gender)) + 
  ylab('Median Friend Count') + 
  xlab('Age')
```

***

### Thinking in Ratios

***

### Wide and Long Format

***

### Reshaping Data

```{r}
#install.packages('reshape2')
#library(reshape2)

pf.fc_by_age_gender.wide <-
  subset(pf.fc_by_age_gender[c('age', 'gender', 'median_friend_count')],
         !is.na(gender)) %>%
  spread(gender, median_friend_count) %>%
  mutate(ratio = male / female)

head(pf.fc_by_age_gender.wide)
```


***

### Ratio Plot

```{r Ratio Plot}
ggplot(data = pf.fc_by_age_gender.wide, 
       mapping = aes(x = age, y = female/male)) +
  geom_line() + ylab('Ratio Females/Males') + xlab('Age') +
  scale_x_continuous(breaks = seq(10,120,5)) + 
  geom_hline(yintercept = 1, linetype = 2)
```

***

### Third Quantitative Variable

```{r Third Quantitative Variable}
pf$year_joined <- floor(2014 - pf$tenure/365)
```

***

### Cut a Variable

```{r Cut a Variable}
sort(table(pf$year_joined))
?cut

pf$year_joined.bucket <-  cut(pf$year_joined,
                              c(2004, 2009, 2011, 2012, 2014))

table(pf$year_joined.bucket)
```

***

### Friending Rate

```{r Friending Rate}
pf <- na.omit(pf)
with(data = subset(pf, tenure >= 1), summary(friend_count/tenure))
```

***

### Friendships Initiated

What is the median friend rate? 0.2209

What is the maximum friend rate? 417

```{r Friendships Initiated}
ggplot(data = subset(pf, tenure >= 1), 
       aes(x = tenure,y = friend_count/tenure)) + 
  geom_line(mapping = aes(color = year_joined.bucket), 
            stat = 'summary', fun.y = mean)
```

***

### Bias-Variance Tradeoff Revisited

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

```

```{r message=FALSE}
ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(mapping = aes(color = year_joined.bucket))
```

***

### Sean's NFL Fan Sentiment Study

***

### Introducing the Yogurt Data Set

```{r Yogurt Data}
yo <- read.csv('./yogurt.csv')

yo$id <- factor(yo$id)
str(yo)
```

***

### Histograms Revisited

```{r Histograms Revisited}
ggplot(data = yo, mapping = aes(x = price)) + 
  geom_histogram() + scale_x_continuous(breaks = seq(0,70,5))
```

***

### Number of Purchases

```{r Number of Purchases}
yo <- transform(yo, all.purchases = strawberry + blueberry + 
                  pina.colada + plain + mixed.berry)
```

***

### Prices over Time

```{r Prices over Time}
ggplot(data = yo, mapping = aes(x = all.purchases)) + 
  geom_histogram() + scale_x_continuous(breaks = seq(0,25,1))

ggplot(data = yo, mapping = aes(x = time, y = price)) + 
  geom_jitter(alpha = 1/4)
```

***

### Sampling Observations

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
set.seed(193264)
sample_ids <- sample(levels(yo$id), 16)

ggplot(data = subset(yo, id %in% sample_ids), 
       mapping = aes(x = time, y = price)) +
  facet_wrap(~id) + 
  geom_line() + 
  geom_point(mapping = aes(size = all.purchases), pch = 1)
```

***

### The Limits of Cross Sectional Data

***

### Many Variables

***

### Scatterplot Matrix
```{r}
#install.packages('GGally')
library(GGally)
```

***

### Even More Variables

***

### Heat Maps

```{r Gene Data}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r Heat Map}
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("orange", "black"))(100))
```

***