Lesson 6
========================================================

***

### Scatterplot Review

```{r Scatterplot Review}
data("diamonds")
library(ggplot2)

ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  scale_x_continuous(limits = c(0,quantile(diamonds$carat,0.99))) +
  scale_y_continuous(limits = c(0,quantile(diamonds$price,0.99))) +
  geom_point(na.rm = TRUE,color = 'black', fill = 'orange', shape = 21)
```

***

### Price and Carat Relationship
Response: Non-linear looks more exponential

***

```{r message=FALSE, warning=FALSE}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  scale_x_continuous(limits = c(0,quantile(diamonds$carat,0.99))) +
  scale_y_continuous(limits = c(0,quantile(diamonds$price,0.99))) +
  geom_point(na.rm = TRUE) + geom_smooth(method = 'lm')
```


#### A diamond is FOREVER


***

### ggpairs Function

```{r ggpairs Function, message=FALSE}
# install these if necessary

# install.packages('GGally')
#install.packages('scales')
#install.packages('memisc')
# install.packages('lattice')
# install.packages('MASS')
# install.packages('car')
# install.packages('reshape')
# install.packages('plyr')

# load the ggplot graphics package and the others
library(ggplot2)
library(GGally)
library(scales)
library(memisc)

# sample 10,000 diamonds from the data set
set.seed(20022012)
diamond_samp <- diamonds[sample(1:length(diamonds$price), 10000), ]
ggpairs(diamond_samp,
  lower = list(continuous = wrap("points", shape = I('.'))),
  upper = list(combo = wrap("box", outlier.shape = I('.'))))
```

***

### The Demand of Diamonds

```{r The Demand of Diamonds, message=FALSE, warning=FALSE}
library(gridExtra)

plot1 <- ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_histogram() +
  ggtitle('Price')

plot2 <- plot1 + scale_x_log10() + ggtitle('Price (log10)')

grid.arrange(plot1,plot2)
```

***

### Scatterplot Transformation

```{r Scatterplot Transformation}
ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_point() + 
  scale_y_log10()
```


### Create a new function to transform the carat variable

```{r cuberoot transformation}
cuberoot_trans = function() trans_new('cuberoot', transform = function(x) x^(1/3),
                                      inverse = function(x) x^3)
```

#### Use the cuberoot_trans function
```{r Use cuberoot_trans}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(na.rm = TRUE) + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Overplotting Revisited

```{r Sort and Head Tables}
head(sort(table(diamonds$carat),decreasing = T))
head(sort(table(diamonds$price),decreasing = T))
```


```{r Overplotting Revisited}
ggplot(aes(carat, price), data = diamonds) + 
  geom_point(na.rm = TRUE, alpha = 1/20, size = 3/4, position = 'jitter') + 
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat')
```

***

### Other Qualitative Factors
Notes: Clarity is another feature which might be useful

***

### Price vs. Carat and Clarity

Alter the code below.
```{r Price vs. Carat and Clarity}
# install and load the RColorBrewer package
# install.packages('RColorBrewer')
library(RColorBrewer)

ggplot(aes(x = carat, y = price), data = diamonds) + 
  geom_point(na.rm = TRUE, alpha = 0.5, size = 1, position = 'jitter',
             mapping = aes(color = clarity)) +
  scale_color_brewer(type = 'div',
    guide = guide_legend(title = 'Clarity', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
    breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
    breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Clarity and Price
Response: Price does vary by clarity, diamonds with higher clarity have higher prices

***

### Price vs. Carat and Cut

Alter the code below.
```{r Price vs. Carat and Cut}
ggplot(aes(x = carat, y = price, color = clarity), data = diamonds) + 
  geom_point(na.rm = TRUE, alpha = 0.5, size = 1, position = 'jitter',
             mapping = aes(color = cut)) +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Clarity', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Clarity')
```

***

### Cut and Price
Response: Not much variation is explained by coloring using the cut of the diamonds

***

### Price vs. Carat and Color

Alter the code below.
```{r Price vs. Carat and Color}
ggplot(data = diamonds, aes(x = carat, y = price, color = color)) + 
  geom_point(na.rm = TRUE, alpha = 0.5, size = 1, position = 'jitter') +
  scale_color_brewer(type = 'div',
                     guide = guide_legend(title = 'Color', reverse = T,
                                          override.aes = list(alpha = 1, size = 2))) +  
  scale_x_continuous(trans = cuberoot_trans(), limits = c(0.2, 3),
                     breaks = c(0.2, 0.5, 1, 2, 3)) + 
  scale_y_continuous(trans = log10_trans(), limits = c(350, 15000),
                     breaks = c(350, 1000, 5000, 10000, 15000)) +
  ggtitle('Price (log10) by Cube-Root of Carat and Cut')
```

***

### Linear Models in R
***

### Building the Linear Model

```{r Building the Linear Model}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamonds)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)
mtable(m1, m2, m3, m4, m5)
```

Notice how adding cut to our model does not help explain much of the variance
in the price of diamonds. This fits with out exploration earlier.

***

### Model Problems

Research:
(Take some time to come up with 2-4 problems for the model)
(You should 10-20 min on this)

***

### A Bigger, Better Data Set

```{r A Bigger, Better Data Set}
#install.package('bitops') # throws an error
#install.packages('RCurl')
#library('bitops')
library('RCurl')

#diamondsurl = getBinaryURL("https://raw.github.com/solomonm/diamonds-data/master/BigDiamonds.Rda")
#load(rawConnection(diamondsurl))
load('./BigDiamonds.Rda')
names(diamondsbig)
```

The code used to obtain the data is available here:
https://github.com/solomonm/diamonds-data

## Building a Model Using the Big Diamonds Data Set

```{r Building a Model Using the Big Diamonds Data Set}
m1 <- lm(I(log(price)) ~ I(carat^(1/3)), data = diamondsbig)
m2 <- update(m1, ~ . + carat)
m3 <- update(m2, ~ . + cut)
m4 <- update(m3, ~ . + color)
m5 <- update(m4, ~ . + clarity)

suppressMessages(library(lattice))
suppressMessages(library(MASS))
suppressMessages(library(memisc))
mtable(m1, m2, m3, m4, m5)
```

***

## Predictions

Example Diamond from BlueNile:
Round 1.00 Very Good I VS1 $5,601

```{r}
# Be sure you’ve loaded the library memisc and have m5 saved as an object in your workspace.
thisDiamond = data.frame(carat = 1.00, cut = "V.Good",
                         color = "I", clarity="VS1")
modelEstimate = predict(m5, newdata = thisDiamond,
                        interval="prediction", level = .95)
```


```{r}
dat = data.frame(m4$model, m4$residuals)

with(dat, sd(m4.residuals))

with(subset(dat, carat > .9 & carat < 1.1), sd(m4.residuals))

dat$resid <- as.numeric(dat$m4.residuals)
ggplot(aes(y = resid, x = round(carat, 2)), data = dat) +
  geom_line(stat = "summary", fun.y = sd)
```


***

